---
title: 'Evens'
author: "Basilis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
df <- read.csv('~/Documents/Langs/Even/apply/scripts/all_morphemes.csv')
library(binom)
library(Hmisc)
library(tidyverse)
```
# morphemes

#### Gre 
(we suppose that it is clearly dialectal)

```{r}
wilcox.test(Gr.E..hab.v/v~corpus, data=df)
U_Gre = binconf(df$Gr.E..hab.v, df$v, return.df = T)$Upper
L_Gre = binconf(df$Gr.E..hab.v, df$v, return.df = T)$Lower
df$speaker <- factor(df$speaker, levels = df$speaker[order(df$Gr.E..hab.v/df$v)])
df %>% ggplot(aes(speaker, Gr.E..hab.v/v, colour=corpus)) + geom_bar(stat="identity") + 
  geom_errorbar(data=df, mapping=aes(ymin=(Gr.E..hab.v-L_Gre*Gr.E..hab.v)/v, ymax=(Gr.E..hab.v+U_Gre*Gr.E..hab.v)/v)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

#### Plural 
(we suppose that it is clearly non-dialectal)

```{r}
wilcox.test(L.pl.n/n~corpus, data=df)

U_L = binconf(df$L.pl.n, df$n, return.df = T)$Upper
L_L = binconf(df$L.pl.n, df$n, return.df = T)$Lower
df$speaker <- factor(df$speaker, levels = df$speaker[order(df$L.pl.n/df$overall)])
df %>% ggplot(aes(speaker, L.pl.n/overall, colour=corpus)) + geom_bar(stat="identity") + geom_errorbar(data=df, mapping=aes(ymin=(L.pl.n-L_L*L.pl.n)/overall, ymax=(L.pl.n+U_L*L.pl.n)/overall)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

####Locative on names 
(we suppose it has some significant difference between the two dialects)

```{r}
wilcox.test(X.dU.LE.loc.n/overall~corpus, data=df)
U = binconf(df$X.dU.LE.loc.n, df$overall, return.df = T)$Upper
L = binconf(df$X.dU.LE.loc.n, df$overall, return.df = T)$Lower
df$speaker <- factor(df$speaker, levels = df$speaker[order(df$X.dU.LE.loc.n/df$overall)])
df %>% ggplot(aes(speaker, X.dU.LE.loc.n/overall, colour=corpus)) + geom_bar(stat="identity") + geom_errorbar(data=df, mapping=aes(ymin=(X.dU.LE.loc.n-L*X.dU.LE.loc.n)/overall, ymax=(X.dU.LE.loc.n+U*X.dU.LE.loc.n)/overall)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

But let's stop and redo the calculations against nouns this time and not all the words, because this is more linguistically reasonable.
#### Locative on names against names
 
```{r}
wilcox.test(X.dU.LE.loc.n/n~corpus, data=df)
U = binconf(df$X.dU.LE.loc.n, df$n, return.df = T)$Upper
L = binconf(df$X.dU.LE.loc.n, df$n, return.df = T)$Lower
df$speaker <- factor(df$speaker, levels = df$speaker[order(df$X.dU.LE.loc.n/df$n)])
df %>% ggplot(aes(speaker, X.dU.LE.loc.n/n, colour=corpus)) + geom_bar(stat="identity") + 
  geom_errorbar(data=df, mapping=aes(ymin=(X.dU.LE.loc.n-L*X.dU.LE.loc.n)/n, ymax=(X.dU.LE.loc.n+U*X.dU.LE.loc.n)/n)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

it gives other results than against all words. Why? Probably because of the undisclosed difference in distribution of nouns. Lets's see
#### nouns distribution

```{r}
wilcox.test(n/overall~corpus, data=df)
U = binconf(df$n, df$overall, return.df = T)$Upper
L = binconf(df$n, df$overall, return.df = T)$Lower
df$speaker <- factor(df$speaker, levels = df$speaker[order(df$n/df$overall)])
df %>% ggplot(aes(speaker, n/overall, colour=corpus)) + geom_bar(stat="identity") + 
  geom_errorbar(data=df, mapping=aes(ymin=(n-L*n)/overall, ymax=(n+U*n)/overall)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


# PCA
PCA when we exclude outliers looks like this

```{r}
library(tidyverse)
even <- read_csv("~/Downloads/together.csv")
even %>% 
  mutate_at(.funs = funs(./overall), 
            .vars = vars(`(E)ldE-soc-v`:`ŋsI-impf.cvb-v`)) %>% 
  mutate(speaker = paste(speaker, corpus)) %>% 
  select(-overall)->
  even

even <- as.data.frame(even)
rownames(even) <- even$speaker
even$speaker <- NULL
corp <- even$corpus
even$corpus <- NULL

pca <- prcomp(even[,-1])
summary(pca)
library(ggfortify)
autoplot(pca,
         shape = FALSE,
         loadings = TRUE,
         label = TRUE,
         loadings.label = TRUE)+
  theme_bw()
```

We suppose the dialects would cluster

because they are different, because they are different in frequency

The hypothesis is that the dialects are clearly clustered because of the difference in frequency 

We need a visualization that shows this

```{r}
corpus_dist=cbind(even, pca$x)

corpus_dist %>% 
  ggplot(aes(PC1, PC2, color=corp))+
  geom_point()+
  stat_ellipse()+
  theme_bw()
```

The dialects are slightly clustered, though outliers are seemingly random (consider the blue dot below to the right). Maybe PCA is a good tool for analyzing our data.
